---
title: 'University of Michigan: Consumer Sentiment (Lagged Linear Models)'
author: 'Group 3: Michelle Nguyen, Jerden Daniels, Cassandra Garner, Eli Hymson, Andrew
  Winn'
output:
word_document: default
pdf_document: default
header-includes: \usepackage{amsmath}
---

```{r}
# Provides data frame 'full' with UMCSENT series and regressors
source(url('https://raw.githubusercontent.com/andrewwinn3/stat626/master/Data/Setup.R'))
attach(full.train)
```

# ARIMA for UMCSENT
This is mostly a duplicate of what we did before. However, we need the UMCSENT series to be stationary, and we will also employ a model which combines the predicted values from the ARIMA UMCSENT model with the covariate series.
```{r}
lm.UMCSENT = auto.arima(UMCSENT, lambda='auto', seasonal=FALSE, stepwise = FALSE)
lm.UMCSENT
checkresiduals(lm.UMCSENT)
UMCSENT.fitted = fitted(lm.UMCSENT)

# I don't know why the a (2,0,2) model isn't recommended here, since UMCSENT.stab is produced by the recommendationded transformation in the previous code, being a (2,1,2) model with $\lambda = 2$ Box-Cox transformation.
lm.UMCSENT.stab = auto.arima(UMCSENT.stab, seasonal=FALSE, stepwise=FALSE)
lm.UMCSENT.stab
checkresiduals(lm.UMCSENT.stab)
UMCSENT.stab.fitted = fitted(lm.UMCSENT.stab)
```


# Models for regressors

### SP:
ARIMA(0,1,0) model with $\delta = 0.007_{(0.002)}$ drift and log transformation.
```{r}
lm.SP = auto.arima(log(SP), seasonal=FALSE, stepwise=FALSE)
lm.SP
checkresiduals(lm.SP)
```

### DGS1:
None of the transformations provide stationarity. Some high-order residual autocorrelations end up being significant.
```{r}
lm.DGS1.log = auto.arima(DGS1.log.diff3[-(1:3)], max.p=10, max.q=10, lambda="auto", seasonal=FALSE, stepwise=FALSE)
lm.DGS1.log
checkresiduals(lm.DGS1.log)
```

### UNRATE:
ARIMA(1,1,3) model with log transformation.
```{r}
lm.UNRATE = auto.arima(log(UNRATE), seasonal=FALSE, stepwise=FALSE)
lm.UNRATE
checkresiduals(lm.UNRATE)
```

### WTI:
ARIMA(2,1,1) model with log transformation.
```{r}
lm.WTI = auto.arima(log(WTI), seasonal=FALSE, stepwise=FALSE)
lm.WTI
checkresiduals(lm.WTI)
```

# Examination of cross-correlation
```{r}
# Stabilized models are denoted by *.stab
ccf(UMCSENT.stab, SP.stab, ylab='CCF')
ccf(UMCSENT.stab, UNRATE.stab, ylab='CCF')
ccf(UMCSENT.stab, WTI.stab, ylab='CCF')
```  

 
# Lagged regression models
These models are based solely on the regressors, rather than any autoregression. (We probably shouldn't include this bit in the report.)
```{r}
lm.leaps = regsubsets(UMCSENT.stab ~ SP.stab + SP.stab.lag1 + SP.stab.lag2 +
                      SP.stab.lag3 + UNRATE.stab.lag1 + WTI.stab.lag1, data = full.train)
summary(lm.leaps)

lm.leaps.1 = lm(UMCSENT.stab ~ SP.stab.lag1)
lm.leaps.2 = lm(UMCSENT.stab ~ SP.stab.lag1 + WTI.stab.lag1)
lm.leaps.3 = lm(UMCSENT.stab ~ SP.stab.lag1 + WTI.stab.lag1 + SP.stab)
lm.leaps.4 = lm(UMCSENT.stab ~ SP.stab.lag1 + WTI.stab.lag1 + SP.stab + SP.stab.lag2)

lm.leaps.AIC = c(
  AIC(lm.leaps.1),
  AIC(lm.leaps.2),
  AIC(lm.leaps.3),
  AIC(lm.leaps.4)
)
lm.leaps.BIC = c(
  BIC(lm.leaps.1),
  BIC(lm.leaps.2),
  BIC(lm.leaps.3),
  BIC(lm.leaps.4)
)

summary(lm.leaps.1)
summary(lm.leaps.2)
summary(lm.leaps.3)
summary(lm.leaps.4)

cat('\nAIC for 1-4 variable models\n', lm.leaps.AIC)
cat('\nBIC for 1-4 variable models\n', lm.leaps.BIC)
cat('\nI recommend four-variable model.')

plot(lm.leaps.4)
```

# Lagged reg. with ARIMA errors
We now consider some lagged regression models which have ARIMA errors. We use backward selection, starting with a four-variable model.

```{r}
lm.arima.4 = auto.arima(log(UMCSENT), seasonal=FALSE, stepwise=FALSE,
                        xreg=cbind(SP.stab.lag1, WTI.stab.lag1,
                                   UNRATE.stab.lag1, SP.stab.lag3))
summary(lm.arima.4)
cat('\nStandardized regression coefficients:\n')
lm.arima.4.stdcoef=lm.arima.4$coef/sqrt(diag(lm.arima.4$var.coef))
lm.arima.4.stdcoef

#Remove SP.stab.lag3
lm.arima.3 = auto.arima(log(UMCSENT), seasonal=FALSE, stepwise=FALSE,
                        xreg=cbind(SP.stab.lag1, WTI.stab.lag1, UNRATE.stab.lag1))
summary(lm.arima.3)
cat('\nStandardized regression coefficients:\n')
lm.arima.3.stdcoef=lm.arima.3$coef/sqrt(diag(lm.arima.3$var.coef))
lm.arima.3.stdcoef

#Remove UNRATE.stab.lag1
lm.arima.2 = auto.arima(log(UMCSENT), seasonal=FALSE, stepwise=FALSE,
                        xreg=cbind(SP.stab.lag1, WTI.stab.lag1))
summary(lm.arima.2)
cat('\nStandardized regression coefficients:\n')
lm.arima.2.stdcoef=lm.arima.2$coef/sqrt(diag(lm.arima.2$var.coef))
lm.arima.2.stdcoef
```

Backward selection brings us to the regressive model with lag 1 (differenced) S&P 500 series, as well as the lag 1 (differenced) WTI series.
