---
title: 'University of Michigan: Consumer Sentiment (Lagged Linear Models)'
author: 'Group 3: Michelle Nguyen, Jerden Daniels, Cassandra Garner, Eli Hymson, Andrew
  Winn'
output:
word_document: default
pdf_document: default
header-includes: \usepackage{amsmath}
---

```{r}
# Provides data frame 'full' with UMCSENT series and regressors
source(url('https://raw.githubusercontent.com/andrewwinn3/stat626/master/Data/Setup.R'))
attach(full.train)
```

# ARIMA for UMCSENT
This is mostly a duplicate of what we did before. However, we need the UMCSENT series to be stationary, and we will also employ a model which combines the predicted values from the ARIMA UMCSENT model with the covariate series.
```{r}
lm.UMCSENT = auto.arima(UMCSENT, lambda='auto', seasonal=FALSE, stepwise = FALSE)
lm.UMCSENT
checkresiduals(lm.UMCSENT)
UMCSENT.fitted = fitted(lm.UMCSENT)

# I don't know why the a (2,0,2) model isn't recommended here, since UMCSENT.stab is produced by the recommendationded transformation in the previous code, being a (2,1,2) model with $\lambda = 2$ Box-Cox transformation.
lm.UMCSENT.stab = auto.arima(UMCSENT.stab, seasonal=FALSE, stepwise=FALSE)
lm.UMCSENT.stab
checkresiduals(lm.UMCSENT.stab)
UMCSENT.stab.fitted = fitted(lm.UMCSENT.stab)
```


# Models for regressors

### SP:
ARIMA(0,1,0) model with $\delta = 0.0089_{(0.0025)}$ drift and $\lambda = 0.041$ B-C transformation.
```{r}
lm.SP = auto.arima(SP, lambda='auto', seasonal=FALSE, stepwise=FALSE)
lm.SP
checkresiduals(lm.SP)
```

### DGS1:
None of the transformations provide stationarity. Some high-order residual autocorrelations end up being significant.
```{r}
lm.DGS1.log = auto.arima(DGS1.log.diff3[-(1:3)], max.p=10, max.q=10, lambda="auto", seasonal=FALSE, stepwise=FALSE)
lm.DGS1.log
checkresiduals(lm.DGS1.log)
```

### UNRATE:
ARIMA(1,1,1) model with $\lambda = -1.000$ B-C transformation.
```{r}
lm.UNRATE = auto.arima(UNRATE, lambda="auto", seasonal=FALSE, stepwise=FALSE)
lm.UNRATE
checkresiduals(lm.UNRATE)
```

### WTI:
ARIMA(1,1,0) model with drift and $\lambda = 0.221$ B-C transformation.
```{r}
lm.WTI = auto.arima(WTI, lambda="auto", seasonal=FALSE, stepwise=FALSE)
lm.WTI
checkresiduals(lm.WTI)
```

# Examination of cross-correlation
```{r}
# Stabilized models are denoted by *.stab
ccf(UMCSENT.stab, SP.stab)
ccf(UMCSENT.stab, UNRATE.stab)
ccf(UMCSENT.stab, WTI.stab)
```

# Lagged regression models
These models are based solely on the regressors, rather than any autoregression.
```{r}
lm.leaps = regsubsets(UMCSENT.stab ~ SP.stab.lag1 + SP.stab.lag3 +
                        UNRATE.stab.lead2 + WTI.stab.lag1, data = full.train)
summary(lm.leaps)

lm.leaps.1 = lm(UMCSENT.stab ~ SP.stab.lag1)
lm.leaps.2 = lm(UMCSENT.stab ~ SP.stab.lag1 + WTI.stab.lag1)
lm.leaps.3 = lm(UMCSENT.stab ~ SP.stab.lag1 + WTI.stab.lag1 + UNRATE.stab.lead2)
lm.leaps.4 = lm(UMCSENT.stab ~ SP.stab.lag1 + WTI.stab.lag1 + UNRATE.stab.lead2 + 
                  SP.stab.lag3)

lm.leaps.AIC = c(
  AIC(lm.leaps.1),
  AIC(lm.leaps.2),
  AIC(lm.leaps.3),
  AIC(lm.leaps.4)
)
lm.leaps.BIC = c(
  BIC(lm.leaps.1),
  BIC(lm.leaps.2),
  BIC(lm.leaps.3),
  BIC(lm.leaps.4)
)

summary(lm.leaps.1)
summary(lm.leaps.2)
summary(lm.leaps.3)
summary(lm.leaps.4)

cat('\nAIC for 1-4 variable models\n', lm.leaps.AIC)
cat('\nBIC for 1-4 variable models\n', lm.leaps.BIC)
cat('\nI recommend four-variable model.')

plot(lm.leaps.4)
```

We now consider some models which utilize the fitted values from the UMCSENT ARIMA model investigated previously.

```{r}
lm.leaps.wFit = regsubsets(UMCSENT.stab ~ UMCSENT.stab.fitted + SP.stab.lag1 + SP.stab.lag3 +
                        UNRATE.stab.lead2 + WTI.stab.lag1, data = full.train)
summary(lm.leaps.wFit)

lm.leaps.wFit.1 = lm(UMCSENT.stab ~ SP.stab.lag1)
lm.leaps.wFit.2 = lm(UMCSENT.stab ~ UMCSENT.stab.fitted + SP.stab.lag1)
lm.leaps.wFit.3 = lm(UMCSENT.stab ~ UMCSENT.stab.fitted + SP.stab.lag1 + WTI.stab.lag1)
lm.leaps.wFit.4 = lm(UMCSENT.stab ~ UMCSENT.stab.fitted + SP.stab.lag1 + WTI.stab.lag1 +
                  UNRATE.stab.lead2)
lm.leaps.wFit.5 = lm(UMCSENT.stab ~ UMCSENT.stab.fitted + SP.stab.lag1 + WTI.stab.lag1 +
                  UNRATE.stab.lead2 + SP.stab.lag3)

lm.leaps.wFit.AIC = c(
  AIC(lm.leaps.wFit.1),
  AIC(lm.leaps.wFit.2),
  AIC(lm.leaps.wFit.3),
  AIC(lm.leaps.wFit.4),
  AIC(lm.leaps.wFit.5)
)
lm.leaps.wFit.BIC = c(
  BIC(lm.leaps.wFit.1),
  BIC(lm.leaps.wFit.2),
  BIC(lm.leaps.wFit.3),
  BIC(lm.leaps.wFit.4),
  BIC(lm.leaps.wFit.5)
)

summary(lm.leaps.wFit.1)
summary(lm.leaps.wFit.2)
summary(lm.leaps.wFit.3)
summary(lm.leaps.wFit.4)
summary(lm.leaps.wFit.5)

cat('\nAIC for 1-4 variable models\n', lm.leaps.wFit.AIC)
cat('\nBIC for 1-4 variable models\n', lm.leaps.wFit.BIC)
cat('\nThe four-variable model is recommended.')

plot(lm.leaps.wFit.4)
```
