---
title: 'University of Michigan: Consumer Sentiment (Report 2)'
author: 'Group 3: Michelle Nguyen, Jerden Daniels, Cassandra Garner, Eli Hymson, Andrew
  Winn'
output:
  word_document: default
  pdf_document: default
header-includes: \usepackage{amsmath}
---

\newcommand{\supt}[1]{^{(\text{#1})}}

```{r setup}

library(tidyverse)
library(readr)
library(TSstudio)
library(xts)
library(quantmod)
library(forecast)
library(tis)
library(gridExtra)
library(timeDate)
library(urca)
```

```{r}
# Functionally identical to Presentation1.Rmd

# 'UMCSENT2.csv' is the FRED file trimed from the first observation of 1978
umcsent <- read_csv(file = '~/Downloads/UMCSENT-2.csv')
umcsent.df <- data.frame(
   DATE <- as.Date(umcsent$DATE),
   UMSCENT = as.numeric(umcsent$UMCSENT))

 UMSCENT = as.numeric(umcsent$UMCSENT)

unrate = getSymbols('UNRATE',src='FRED', auto.assign = F, from = "1978-01-01" , to = "2020-05-01") 
unrate.df = data.frame(date=time(unrate), coredata(unrate))
recessions.df = read.table(textConnection(
"Peak, Trough
1873-10-01, 1879-03-01
1882-03-01, 1885-05-01
1887-03-01, 1888-04-01
1890-07-01, 1891-05-01
1893-01-01, 1894-06-01
1895-12-01, 1897-06-01
1899-06-01, 1900-12-01
1902-09-01, 1904-08-01
1907-05-01, 1908-06-01
1910-01-01, 1912-01-01
1913-01-01, 1914-12-01
1918-08-01, 1919-03-01
1920-01-01, 1921-07-01
1923-05-01, 1924-07-01
1926-10-01, 1927-11-01
1929-08-01, 1933-03-01
1937-05-01, 1938-06-01
1945-02-01, 1945-10-01
1948-11-01, 1949-10-01
1953-07-01, 1954-05-01
1957-08-01, 1958-04-01
1960-04-01, 1961-02-01
1969-12-01, 1970-11-01
1973-11-01, 1975-03-01
1980-01-01, 1980-07-01
1981-07-01, 1982-11-01
1990-07-01, 1991-03-01
2001-03-01, 2001-11-01
2007-12-01, 2009-06-01"), sep=',',
colClasses=c('Date', 'Date'), header=TRUE)

recessions.trim = subset(recessions.df, Peak >= min(unrate.df$date) )
Yt = ts(umcsent$UMCSENT, frequency = 12, start = c(1978,1))

# UMCSENT separated by components
comp = read.csv("~/Desktop/school/STAT 626/UMCSENT_COMPONENTS.csv")
comp$DATE = as.Date(paste(comp$Year, comp$Month, "01", sep="-"))
```

```{r}
# Perform lag 1 differencing
umcsent.df = data.frame(umcsent.df, LAG1 = c(0, diff(umcsent$UMCSENT,1)))
colnames(umcsent.df)[1] = "DATE"

ggplot(data=umcsent.df) +
  geom_line(aes(x=DATE, y=LAG1))

# Residual plots suggest normality, homosketasticity, and constant mean. (see below)
# Moreover, the intercept of the differenced series is not significant, suggesting
#    that the original series has no general trend.
lag1lm = lm(LAG1 ~ DATE, data = umcsent.df)
plot(lag1lm)
summary(lag1lm)

# The Dickey-Fuller test supports stationarity of lag 1 transformation at p < 0.01
aTSA::adf.test(umcsent.df$LAG1)
```  
The at first glance the residuals seem to have approximately constant variance. However, we see that there are regions with higher error variance, which corresponds to increased volatility in the original series. These regions are certainly worth noting, but the nonuniformity of noise variances is minor enough that it is likely permissible to ignore.

```{r}
# Repeat these procedures with comp, the dataframe of component series.
# Abbreviations:
# PFC: current personal finance conditions
# PFE: personal finance expectations (12 months)
# BC12: business conditions (12 months)
# BC5: business conditions (5 years)
# BC: current buying conditions
comp$PFC.LAG1 = c(0, diff(comp$Personal.Finance.Current,1))
comp$PFE.LAG1 = c(0, diff(comp$Personal.Finance.Expected,1))
comp$BC12.LAG1 = c(0, diff(comp$Business.Condition.12.Months,1))
comp$BC5.LAG1 = c(0, diff(comp$Business.Condition.5.Years,1))
comp$BC.LAG1 = c(0, diff(comp$Buying.Conditions,1))

# Restructures comp dataframe to allow for use of faceting
comp.facet = data.frame(DATE=rep(comp$DATE, 5),
                        COMPONENT=rep(c('PFC', 'PFE', 'BC12', 'BC5', 'BC'), each=nrow(comp)),
                        INDEX=c(comp$PFC.LAG1, comp$PFE.LAG1, comp$BC12.LAG1, comp$BC5.LAG1, comp$BC.LAG1))

# Plots lag 1 differencing of components
ggplot(data=comp.facet) + 
  geom_line(aes(x=DATE, y=INDEX)) + 
  facet_grid(COMPONENT~.)

# Check residuals and slope estimates when viewing series as linear models.
# Note that since components take strictly integer values, we'll have some natural
#    clustering, but that is okay.
PFC.lm = lm(PFC.LAG1 ~ DATE, data=comp)
PFE.lm = lm(PFE.LAG1 ~ DATE, data=comp)
BC12.lm = lm(BC12.LAG1 ~ DATE, data=comp)
BC5.lm = lm(BC5.LAG1 ~ DATE, data=comp)
BC.lm = lm(BC.LAG1 ~ DATE, data=comp)

summary(PFC.lm)
plot(PFC.lm)

summary(PFE.lm)
plot(PFE.lm)

summary(BC12.lm)
plot(BC12.lm)

summary(BC5.lm)
plot(BC5.lm)

summary(BC.lm)
plot(BC.lm)

# Dickey-Fuller tests suggest alternative hypothesis of stationarity at p<0.01
aTSA::adf.test(comp$PFC.LAG1)
aTSA::adf.test(comp$PFE.LAG1)
aTSA::adf.test(comp$BC12.LAG1)
aTSA::adf.test(comp$BC5.LAG1)
aTSA::adf.test(comp$BC.LAG1)
```  

The reasonable conclusions for individual components are essentially the same as for the UMCSENT index as a whole. We have the desired normality and regularity assumptions excpet for some (arguably insignificant) regions of volatility in which noise variance increases slightly. (Note that this seems most extreme in the 'buying conditions' series, so any further investigation should begin there. That is also the series for which the assumption of constant mean had the least evidence, although the evidence is still sufficiently strong.) The Dickey-Fuller tests for non-stationarity all return p-values less than 0.01, so it is reasonable to proceed under the assumptions that the lag 1 difference series are stationary.

\section*{Random walk models}
For the lag 1 differenced series, we have already rejected the null hypotheses (at any reasonable $\alpha$-level) that each series is nonstationary. This implies that each of the original, undifferenced time series can be modeled by a random walk with i.i.d. white noise terms. We now wish to determine the presence or absence of any drift or trend.

We previously used some simple linear models as a convenient way of investigating the
white noise variance of each differenced series. Were the original series to have a trend,
we would expect the DATE variable to have a statistically significant coefficient. However,
our linear model diagnostics give us that the most significant DATE coefficient occurs
in the model for the Buying Conditions lag 1 series, where the coefficient is -4.950e-5
with standard error 7.642e-5 (505 degrees of freedom). This is hardly sufficient evidence 
to justify inclusion of a trend term, so we now turn our attention to drift.

Drift will be present in the undifferenced random walks if and only if the lag 1
differenced series have nonzero mean, so we perform the classical t-tests for zero means:

```{r}
t.test(umcsent.df$LAG1)
t.test(comp$PFC.LAG1)
t.test(comp$PFE.LAG1)
t.test(comp$BC12.LAG1)
t.test(comp$BC5.LAG1)
t.test(comp$BC.LAG1)
```

Once again, the Buying Conditions series yields the most significant test statistic. However,
as the 95-percent confidence interval for mean lag-1 difference is $(-0.779, 0.558)$, we
do not reject any null hypotheses that the differenced series have zero means, and as such
we will not include drift in our models. After calculating the empirical variances for the lag 1 differenced series, we have the following simple random walk models:

```{r}
var(umcsent.df$LAG1, na.rm=TRUE)
var(comp$PFC.LAG1, na.rm=TRUE)
var(comp$PFE.LAG1, na.rm=TRUE)
var(comp$BC12.LAG1, na.rm=TRUE)
var(comp$BC5.LAG1, na.rm=TRUE)
var(comp$BC.LAG1, na.rm=TRUE)
```




```{r}

source('~/Downloads/ARMAroots_RFunctions.R')
#ACF AND PACF PLOTS
ggtsdisplay(umcsent.df$LAG1, xlab = "Time", ylab = "index 1966:Q1 = 100", main= "Consumer Sentiment")
ggtsdisplay(comp$PFC.LAG1, xlab = "Time", ylab = "index 1966:Q1 = 100", main= "Current Personal Finance")
ggtsdisplay(comp$PFE.LAG1, xlab = "Time", ylab = "index 1966:Q1 = 100", main= "Expected Personal Finance")
ggtsdisplay(comp$BC12.LAG1, xlab = "Time", ylab = "index 1966:Q1 = 100", main= "Business conditions in 12 months")
ggtsdisplay(comp$BC5.LAG1, xlab = "Time", ylab = "index 1966:Q1 = 100", main= "Business Conditions in 5 Years")
ggtsdisplay(comp$BC.LAG1, xlab = "Time", ylab = "index 1966:Q1 = 100", main= "Buying Conditions")


```

```{r}
#combine all the TS


dTS = data.frame(umcsent.df$LAG1, comp$PFC.LAG1, comp$PFE.LAG1, comp$BC12.LAG1, comp$BC5.LAG1, comp$BC.LAG1)

```


```{r}

n = dim(dTS)[1]
k = dim(dTS)[2]
p.max = trunc(12*(n/100)^.25); cat("Maximum lag for m-AIC and m-BIC:", p.max)

dADF = matrix(ncol = 7, nrow = k)
types = c("drift", "drift", "drift", "none", "none", "none")
df0 = c(1,1,1,1,1,1)

colnames(dADF) = c("cri value", "adf-aic", "lag-aic", "signif-aic", "adf-bic", "lag-bic", "sign-bic")
rownames(dADF) = names(dTS)

for(i in 1:k){
  obj = ur.df(dTS[,i], type = types[i], lag = p.max, selectlags = "AIC");
  dADF[i,1] = obj@cval[1,2]
  dADF[i,2] = obj@teststat[1]
  dADF[i,3] = obj@testreg$df[1] - df0[i]
  dADF[i,4] = dADF[i,2] < dADF[i,1] #1: Reject Ho
  obj = ur.df(dTS[,i], type = types[i], lag = p.max, selectlags = "BIC");
  dADF[i,5] = obj@teststat[1]
  dADF[i,6] = obj@testreg$df[1] - df0[i]
  dADF[i,7] = dADF[i,5] < dADF[i,1] #1: Reject Ho
  
}
cat("\nADF tests of differenced series\n"); dADF


```

```{r}
#fitting the TS with AIC

fit1aic = auto.arima(dTS[,1], d = 1, seasonal = F, ic = "aic")
DF1 = round(log(length(dTS[,1])))
fit2aic = auto.arima(dTS[,2], d = 1, seasonal = F, ic = "aic")
DF2 = round(log(length(dTS[,2])))
fit3aic = auto.arima(dTS[,3], d = 1, seasonal = F, ic = "aic")
DF3 = round(log(length(dTS[,3])))
fit4aic = auto.arima(dTS[,4], d = 1, seasonal = F, ic = "aic")
DF4 = round(log(length(dTS[,4])))
fit5aic = auto.arima(dTS[,5], d = 1, seasonal = F, ic = "aic")
DF5 = round(log(length(dTS[,5])))
fit6aic = auto.arima(dTS[,6], d = 1, seasonal = F, ic = "aic")
DF6 = round(log(length(dTS[,6])))

#fitting the TS with BIC

fit1bic = auto.arima(dTS[,1], d = 1, seasonal = F, ic = "bic")
fit2bic = auto.arima(dTS[,2], d = 1, seasonal = F, ic = "bic")
fit3bic = auto.arima(dTS[,3], d = 1, seasonal = F, ic = "bic")
fit4bic = auto.arima(dTS[,4], d = 1, seasonal = F, ic = "bic")
fit5bic = auto.arima(dTS[,5], d = 1, seasonal = F, ic = "bic")
fit6bic = auto.arima(dTS[,6], d = 1, seasonal = F, ic = "bic")


```

```{r}
cat("\nModel select by the AIC for UMCSENT:\n") 
fit1aic
cat("\nModel select by the BIC for UMCSENT:\n") 
fit1bic

#Both the AIC and BIC picked the model ARIMA(1,1,0) with drift for UMCSENT

cat("\nModel select by the AIC for PFC:\n") 
fit2aic
cat("\nModel select by the BIC for PFC:\n") 
fit2bic

#Both the AIC and BIC picked the model ARIMA(5,1,0) with drift for PFC

cat("\nModel select by the AIC for PFE:\n") 
fit3aic
cat("\nModel select by the BIC for PFE:\n") 
fit3bic

#Both the AIC and BIC picked the model ARIMA(1,1,0) with drift for PFE

cat("\nModel select by the AIC for BC12:\n") 
fit4aic
cat("\nModel select by the BIC for BC12:\n") 
fit4bic

#Both the AIC and BIC picked the model ARIMA(1,1,0) with drift for BC12. 

cat("\nModel select by the AIC for BC5:\n") 
fit5aic
cat("\nModel select by the BIC for BC5:\n") 
fit5bic

#The AIC picked ARIMA(5,1,0) and the BIC picked ARIMA(4,1,0). Since the lowest AIC is ARIMA(5,1,0), we will choose that model for BC5.

cat("\nModel select by the AIC for BC:\n") 
fit6aic
cat("\nModel select by the BIC for BC:\n") 
fit6bic

#Both the AIC and BIC picked the model ARIMA(5,1,0) with drift for BC.


```

```{r}

cat("Portmanteau tests of fitted residuals")
checkresiduals(fit1aic, lag = DF1 + length(fit1aic$model$phi))
checkresiduals(fit2aic, lag = DF2 + length(fit2aic$model$phi))
checkresiduals(fit3aic, lag = DF3 + length(fit3aic$model$phi))
checkresiduals(fit4aic, lag = DF4 + length(fit4aic$model$phi))
checkresiduals(fit5aic, lag = DF5 + length(fit5aic$model$phi))
checkresiduals(fit6aic, lag = DF6 + length(fit6aic$model$phi))
```
```{r}
#UMSCENT redone

umcsent_diff = diff(UMSCENT)

fitaic = auto.arima(umcsent_diff, d = 1, seasonal = F, ic = "aic")
DF1 = round(log(length(dTS[,1])))
fitbic = auto.arima(dTS[,1], d = 1, seasonal = F, ic = "bic")
fitaic
fitbic
checkresiduals(fitaic, lag = DF1 + length(fitaic$model$phi))

#We choose ARIMA(5,1,0) since it has the lowest AIC and BIC.
#After checking the residuals, the Ljung Box Test's pvalue = 0.00174. We fail to reject the null hypothesis that autocorrelations up to lag 11 equal 0 and conclude the data values are significantly random and independent up to lag 11.
```


\begin{align*}
x\supt{UMCSENT}_t &= x\supt{UMCSENT}_{t-1} + w\supt{UMCSENT}_t&
  w\supt{UMCSENT}_t &\sim N(0, 15.8)\\
x\supt{PFC}_t &= x\supt{PFC}_{t-1} + w\supt{PFC}_t&
  w\supt{PFC}_t &\sim N(0, 38.6)\\
x\supt{PFE}_t &= x\supt{PFE}_{t-1} + w\supt{PFE}_t&
  w\supt{PFE}_t &\sim N(0, 25.5)\\
x\supt{BC12}_t &= x\supt{BC12}_{t-1} + w\supt{BC12}_t&
  w\supt{PBC12}_t &\sim N(0, 117)\\
x\supt{BC5}_t &= x\supt{BC5}_{t-1} + w\supt{BC5}_t&
  w\supt{BC5}_t &\sim N(0, 59.7)\\
x\supt{BC}_t &= x\supt{BC}_{t-1} + w\supt{BC}_t&
  w\supt{BC}_t &\sim N(0, 58.7)\\
\end{align*}
\begin{quote}
UMCSENT: University of Michigan Consumer Sentiment Index (the primary index)\\
PFC: Current personal finance conditions\\
PFE: Expected personal finance conditions (12 months)\\
BC12: Expected business conditions (12 months)\\
BC5: Expected business conditions (5 years)\\
BC: Current buying conditions
\end{quote}